---
title: "R Package Validation Report"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(jsonlite)
library(knitr)

# system_info <- fromJSON("system-info.json")  # temporarily disabled
classification <- fromJSON("package-classification.json")
validation <- fromJSON("validation-results.json", simplifyDataFrame = FALSE)
riskmetric_res <- fromJSON("riskmetric-results.json", simplifyDataFrame = FALSE)
levels_data <- fromJSON("level-assignments.json", simplifyDataFrame = FALSE)
vuln_report <- tryCatch(read.csv("vulnerability-report.csv", stringsAsFactors = FALSE),
                        error = function(e) data.frame())

# Helper to extract level counts
level_vec <- vapply(levels_data, function(x) x$level, integer(1))
```

# Executive Summary

| Item | Value |
|------|-------|
| **Report Date** | `r format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")` |
| **Git Commit** | `r tryCatch(trimws(system("git rev-parse HEAD", intern = TRUE)), error = function(e) "N/A")` |
| **R Version** | `r R.version.string` |
| **Platform** | `r R.version$platform` |
| **Level 1 (Base)** | `r sum(level_vec == 1)` packages |
| **Level 2 (Auto-qualified)** | `r sum(level_vec == 2)` packages |
| **Level 3 (Manual review)** | `r sum(level_vec == 3)` packages |
| **Vulnerabilities** | `r nrow(vuln_report)` findings |

<!-- System Qualification section temporarily disabled (system-info.json BOM issue on Windows)

# System Qualification

## R Environment

- **R Version**: N/A
- **Platform**: N/A
- **OS**: N/A
- **.NET SDK**: N/A
- **Locale**: N/A

## Library Paths

```{r eval=FALSE}
cat(paste(system_info$lib_paths, collapse = "\n"))
```

-->

# Package Classification

## Intended-for-Use (`r length(classification$intended_for_use)` packages)

These packages are directly used in analyses and undergo full validation (R CMD check + coverage + riskmetric).

```{r}
kable(data.frame(Package = classification$intended_for_use))
```

## Import Packages (`r length(classification$imports)` packages)

These are transitive dependencies. They receive riskmetric scoring but not full R CMD check or coverage analysis.

```{r}
if (length(classification$imports) > 0) {
  kable(data.frame(Package = classification$imports))
} else {
  cat("No import packages identified.")
}
```

# Per-Package Validation Results

Detailed results for each intended-for-use package.

```{r results='asis'}
for (pkg_name in classification$intended_for_use) {
  cat("\n## ", pkg_name, "\n\n")

  v <- validation[[pkg_name]]
  r <- riskmetric_res[[pkg_name]]
  l <- levels_data[[pkg_name]]

  cat("| Attribute | Value |\n")
  cat("|-----------|-------|\n")
  cat("| **Source** | ", if (!is.null(v$source)) v$source else "N/A", " |\n")
  cat("| **Git SHA** | ", if (!is.null(v$git_sha) && !is.na(v$git_sha)) v$git_sha else "N/A", " |\n")
  cat("| **R CMD check** | ", if (!is.null(v$rcmdcheck)) v$rcmdcheck else "N/A", " |\n")
  cat("| **Test Coverage** | ",
      if (!is.null(v$coverage_pct) && !is.na(v$coverage_pct)) paste0(round(v$coverage_pct, 1), "%") else "N/A",
      " |\n")
  cat("| **Riskmetric Score** | ",
      if (!is.null(r$riskmetric_score) && !is.na(r$riskmetric_score)) round(r$riskmetric_score, 4) else "N/A",
      " |\n")
  cat("| **Assigned Level** | ",
      if (!is.null(l$level)) l$level else "N/A",
      " |\n")
  cat("| **Level Reason** | ",
      if (!is.null(l$reason)) l$reason else "N/A",
      " |\n")
  cat("\n")

  # R CMD check details
  if (!is.null(v$rcmdcheck_errors) && length(v$rcmdcheck_errors) > 0) {
    cat("**Errors:**\n\n```\n", paste(v$rcmdcheck_errors, collapse = "\n"), "\n```\n\n")
  }
  if (!is.null(v$rcmdcheck_warnings) && length(v$rcmdcheck_warnings) > 0) {
    cat("**Warnings:**\n\n```\n", paste(v$rcmdcheck_warnings, collapse = "\n"), "\n```\n\n")
  }
}
```

# Import Packages Summary

Riskmetric scores for import (transitive dependency) packages.

```{r}
import_rows <- lapply(classification$imports, function(pkg) {
  r <- riskmetric_res[[pkg]]
  l <- levels_data[[pkg]]
  data.frame(
    Package = pkg,
    Riskmetric_Score = if (!is.null(r$riskmetric_score) && !is.na(r$riskmetric_score)) round(r$riskmetric_score, 4) else NA,
    Level = if (!is.null(l$level)) l$level else NA,
    stringsAsFactors = FALSE
  )
})

if (length(import_rows) > 0) {
  import_df <- do.call(rbind, import_rows)
  kable(import_df, row.names = FALSE)
} else {
  cat("No import packages to display.")
}
```

# Vulnerability Scan Results

```{r}
if (nrow(vuln_report) > 0) {
  kable(vuln_report)
} else {
  cat("No vulnerabilities detected.")
}
```

# Session Info

```{r}
sessionInfo()
```

# Reproducibility Manifest

```{r}
manifest_rows <- lapply(classification$intended_for_use, function(pkg) {
  v <- validation[[pkg]]
  data.frame(
    Package = pkg,
    Source = if (!is.null(v$source)) v$source else "N/A",
    Git_SHA = if (!is.null(v$git_sha) && !is.na(v$git_sha)) v$git_sha else "N/A",
    stringsAsFactors = FALSE
  )
})

if (length(manifest_rows) > 0) {
  manifest_df <- do.call(rbind, manifest_rows)
  kable(manifest_df, row.names = FALSE)
} else {
  cat("No manifest entries.")
}
```
