---
title: "R Package Validation Report"
subtitle: "ESQ Cloud Validation Environment"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    latex_engine: xelatex
geometry: "left=3cm,right=2.5cm,top=2.5cm,bottom=2.5cm"
fontsize: 11pt
header-includes:
  - \usepackage{fancyhdr}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{hyperref}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[L]{\small R Package Validation Report}
  - \fancyhead[R]{\small\thepage}
  - \fancyfoot[C]{\small Confidential}
  - \renewcommand{\headrulewidth}{0.4pt}
  - \renewcommand{\footrulewidth}{0.4pt}
  - \fancypagestyle{plain}{\fancyhf{}\fancyfoot[C]{\small Confidential}\renewcommand{\headrulewidth}{0pt}\renewcommand{\footrulewidth}{0.4pt}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(jsonlite)
library(knitr)

classification  <- fromJSON("package-classification.json")
validation      <- fromJSON("validation-results.json",   simplifyDataFrame = FALSE)
riskmetric_res  <- fromJSON("riskmetric-results.json",   simplifyDataFrame = FALSE)
levels_data     <- fromJSON("level-assignments.json",    simplifyDataFrame = FALSE)
vuln_report     <- tryCatch(
  read.csv("vulnerability-report.csv", stringsAsFactors = FALSE),
  error = function(e) data.frame()
)

level_vec <- vapply(levels_data, function(x) x$level, integer(1))

# Helper: safely extract a numeric value, return NA otherwise
safe_num <- function(x) {
  v <- suppressWarnings(tryCatch(as.numeric(x), error = function(e) NA_real_))
  if (length(v) == 1L && !is.na(v)) v else NA_real_
}

# Helper: wrap long lines so verbatim blocks never overflow the page.
# Preserves existing newlines; wraps individual lines that exceed `width`.
# Uses strwrap() for word-boundary wrapping where possible; lines with no
# whitespace (e.g. long paths) are hard-broken with a continuation indent.
wrap_verbatim <- function(text, width = 88) {
  lines <- strsplit(as.character(text), "\n", fixed = TRUE)[[1]]
  result <- unlist(lapply(lines, function(l) {
    if (nchar(l) <= width) return(l)
    wrapped <- strwrap(l, width = width, exdent = 2)
    if (length(wrapped) == 1L && nchar(wrapped) > width) {
      # No whitespace to break on — hard-break every `width` chars
      chunks <- character(0)
      while (nchar(l) > 0) {
        chunks <- c(chunks, substr(l, 1, width))
        l <- substr(l, width + 1L, nchar(l))
      }
      return(chunks)
    }
    wrapped
  }))
  paste(result, collapse = "\n")
}

# Helper: truncate long strings for table cells
trunc_str <- function(s, n = 70) {
  if (is.null(s) || is.na(s)) return("N/A")
  s <- as.character(s)
  if (nchar(s) > n) paste0(substr(s, 1, n - 3), "...") else s
}

# Helper: get package source — prefer validation result, fall back to
# classification refs (guaranteed present for directly listed packages)
pkg_source <- function(pkg, v) {
  if (!is.null(v) && !is.null(v$source) && !is.na(v$source)) return(v$source)
  ref <- classification$intended_for_use_refs[[pkg]]
  if (!is.null(ref) && length(ref) > 0 && !is.na(ref)) ref else "N/A"
}
```

\newpage

# Executive Summary

```{r exec-summary}
git_commit <- tryCatch(
  trimws(system("git rev-parse HEAD", intern = TRUE)),
  error = function(e) "N/A"
)

summary_df <- data.frame(
  Item = c(
    "Report Date",
    "Git Commit",
    "R Version",
    "Platform",
    "Level 1 (Base/Recommended)",
    "Level 2 (Auto-qualified)",
    "Level 3 (Manual review required)",
    "Vulnerabilities"
  ),
  Value = c(
    format(Sys.time(), "%Y-%m-%d %H:%M:%S UTC"),
    git_commit,
    R.version.string,
    R.version$platform,
    paste(sum(level_vec == 1), "packages"),
    paste(sum(level_vec == 2), "packages"),
    paste(sum(level_vec == 3), "packages"),
    paste(nrow(vuln_report), "findings")
  ),
  stringsAsFactors = FALSE
)

kable(summary_df, format = "latex", booktabs = TRUE,
      col.names = c("Item", "Value"), linesep = "") |>
  kableExtra::column_spec(1, bold = TRUE, width = "5cm") |>
  kableExtra::column_spec(2, width = "9cm")
```

# Regulatory Framework

This report supports regulatory compliance for the use of R as a statistical computing environment in pharmaceutical drug development.
The validation approach is consistent with:

- **FDA 21 CFR Part 11** — R is classified as non-transactional software; electronic records requirements apply to data systems, not to the statistical software engine itself.
- **FDA Statistical Software Clarifying Statement (2015)** — The FDA does not mandate specific software. However, documentation of software packages (name, version, build) is required upon submission.
- **R Foundation Guidance** — The R Foundation for Statistical Computing provides official guidance on regulatory compliance. As stated in *A Risk-Based Approach for Assessing R Package Accuracy within a Validated Infrastructure*: "there is minimal risk in using base and recommended (core) packages as a component in a validated system for regulatory analysis and reporting with R."
- **ISO 9001** — Quality management principles are applied to the validation process.

## Validation Methodology

A risk-based 7-step validation process is applied to each package:

1. **Investigation** — Assess intended use, maturity, and quality; define functional requirements.
2. **Investigation Review** — Confirm requirements cover all essential functionality.
3. **Remediation** — Execute unit tests verifying the package satisfies stated requirements.
4. **Remediation Review** — Peer review of test scripts to confirm requirements are met.
5. **Final Review** — Independent check against validation best practices.
6. **Approval** — Final verdict on the package validation status.
7. **Maintenance** — Continuous maintenance of the testing environment and deployment architecture.

## Package Classification

Two categories are applied within this validated installation:

- **Intended for Use** — Packages loaded directly by analysts during an R session. Full validation is applied: R CMD check, test coverage measurement, and riskmetric scoring.
- **Imports** — Transitive dependencies required for Intended-for-Use packages to function. Risk-assessed via riskmetric only; not directly loaded by users.
- **Base/Recommended** — Distributed with R. Classified as Level 1 with no further validation required per R Foundation guidance.

## Validation Levels

| Level | Description | Criteria |
|-------|-------------|----------|
| **1** | Base/Recommended | Distributed with R |
| **2** | Auto-qualified | Intended-for-Use: coverage ≥ `r tryCatch({y <- yaml::read_yaml("validation-thresholds.yml"); paste0(y$coverage_threshold, "%")}, error=function(e)"50%")` AND riskmetric ≤ `r tryCatch({y <- yaml::read_yaml("validation-thresholds.yml"); y$riskmetric_threshold}, error=function(e)"0.5")`; Imports: riskmetric ≤ threshold |
| **3** | Manual review | Criteria not met; explicit risk assessment required |

\newpage

# System Environment

```{r system-info}
sys_df <- data.frame(
  Component = c("R Version", "Platform", "R Library Path"),
  Value     = c(
    R.version.string,
    paste(R.version$platform, R.version$os, sep = " / "),
    trunc_str(.libPaths()[1], 80)
  ),
  stringsAsFactors = FALSE
)

kable(sys_df, format = "latex", booktabs = TRUE,
      col.names = c("Component", "Value"), linesep = "") |>
  kableExtra::column_spec(1, bold = TRUE, width = "4cm") |>
  kableExtra::column_spec(2, width = "10cm")
```

Session information is provided in full in Appendix A.

\newpage

# Package Classification

## Intended-for-Use Packages

`r length(classification$intended_for_use)` package(s) were identified for full validation.

```{r ifu-table}
if (length(classification$intended_for_use) > 0) {
  kable(
    data.frame(Package = classification$intended_for_use, stringsAsFactors = FALSE),
    format = "latex", booktabs = TRUE, linesep = ""
  )
} else {
  cat("No intended-for-use packages identified.")
}
```

## Import Packages

`r length(classification$imports)` transitive dependenc`r if (length(classification$imports) == 1) "y" else "ies"` identified. These are not directly loaded by analysts and receive riskmetric scoring only.

```{r imports-table}
if (length(classification$imports) > 0) {
  n_cols <- 3
  pkgs   <- classification$imports
  pad    <- (n_cols - length(pkgs) %% n_cols) %% n_cols
  padded <- c(pkgs, rep("", pad))
  mat    <- matrix(padded, ncol = n_cols, byrow = TRUE)
  kable(mat, format = "latex", booktabs = TRUE,
        col.names = rep("Package", n_cols), linesep = "")
} else {
  cat("No import packages identified.")
}
```

## Base Packages

`r length(classification$base)` base/recommended package(s) are classified as Level 1 and require no further validation.

\newpage

# Per-Package Validation Results

```{r pkg-results, results='asis'}
thresholds <- tryCatch(yaml::read_yaml("validation-thresholds.yml"),
                       error = function(e) list(coverage_threshold = 50, riskmetric_threshold = 0.5))

for (pkg_name in classification$intended_for_use) {
  cat("\n## ", pkg_name, "\n\n")

  v <- validation[[pkg_name]]
  r <- riskmetric_res[[pkg_name]]
  l <- levels_data[[pkg_name]]

  coverage_pct  <- safe_num(v$coverage_pct)
  risk_score    <- safe_num(r$riskmetric_score)
  rcmdcheck_val <- if (!is.null(v$rcmdcheck)) v$rcmdcheck else "N/A"

  # Determine Pass/Fail indicators
  cov_flag  <- if (!is.na(coverage_pct))
    if (coverage_pct >= thresholds$coverage_threshold) "PASS" else "FAIL" else "N/A"
  risk_flag <- if (!is.na(risk_score))
    if (risk_score <= thresholds$riskmetric_threshold) "PASS" else "FAIL" else "N/A"
  level_val <- if (!is.null(l$level)) l$level else "N/A"

  metrics_df <- data.frame(
    Attribute = c("Source", "Git SHA", "R CMD Check",
                  "Test Coverage", "Coverage Threshold",
                  "Riskmetric Score", "Riskmetric Threshold",
                  "Assigned Level"),
    Value = c(
      trunc_str(pkg_source(pkg_name, v), 65),
      if (!is.null(v$git_sha) && !is.na(v$git_sha)) substr(v$git_sha, 1, 12) else "N/A",
      rcmdcheck_val,
      if (!is.na(coverage_pct)) paste0(round(coverage_pct, 1), "% [", cov_flag, "]")  else "N/A",
      paste0(thresholds$coverage_threshold, "%"),
      if (!is.na(risk_score)) paste0(round(risk_score, 4), " [", risk_flag, "]") else "N/A",
      as.character(thresholds$riskmetric_threshold),
      paste0("Level ", level_val)
    ),
    stringsAsFactors = FALSE
  )

  print(
    kable(metrics_df, format = "latex", booktabs = TRUE,
          col.names = c("Attribute", "Value"), linesep = "") |>
      kableExtra::column_spec(1, bold = TRUE, width = "4.5cm") |>
      kableExtra::column_spec(2, width = "9.5cm")
  )

  cat("\n**Level Reason:** ", trunc_str(l$reason, 110), "\n\n")

  has_errors   <- !is.null(v$rcmdcheck_errors)   && length(v$rcmdcheck_errors)   > 0
  has_warnings <- !is.null(v$rcmdcheck_warnings) && length(v$rcmdcheck_warnings) > 0

  if (has_errors) {
    cat("\n**R CMD Check Errors:**\n\n")
    for (err in v$rcmdcheck_errors) {
      cat("\\begingroup\\small\n\\begin{verbatim}\n")
      cat(wrap_verbatim(err))
      cat("\n\\end{verbatim}\n\\endgroup\n\n")
    }
  }

  if (has_warnings) {
    cat("\n**R CMD Check Warnings:**\n\n")
    for (w in v$rcmdcheck_warnings) {
      cat("\\begingroup\\small\n\\begin{verbatim}\n")
      cat(wrap_verbatim(w))
      cat("\n\\end{verbatim}\n\\endgroup\n\n")
    }
  }

  cat("\\vspace{0.5em}\n\n")
}
```

\newpage

# Import Package Risk Assessment

Risk scores for all transitive dependency packages. Packages with riskmetric score above `r thresholds$riskmetric_threshold` are flagged for review.

```{r import-risk}
if (length(classification$imports) > 0) {
  import_rows <- lapply(classification$imports, function(pkg) {
    r <- riskmetric_res[[pkg]]
    l <- levels_data[[pkg]]
    rs <- safe_num(r$riskmetric_score)
    data.frame(
      Package       = pkg,
      Riskmetric    = if (!is.na(rs)) round(rs, 4) else NA_real_,
      Level         = if (!is.null(l$level)) l$level else NA_integer_,
      stringsAsFactors = FALSE
    )
  })
  import_df <- do.call(rbind, import_rows)
  import_df <- import_df[order(import_df$Package), ]

  kable(import_df, format = "latex", booktabs = TRUE, longtable = TRUE,
        row.names = FALSE,
        col.names = c("Package", "Riskmetric Score", "Level"),
        linesep = "") |>
    kableExtra::column_spec(1, width = "7cm") |>
    kableExtra::column_spec(2, width = "4cm") |>
    kableExtra::column_spec(3, width = "2cm")
} else {
  cat("No import packages to display.")
}
```

\newpage

# Vulnerability Assessment

Installed packages were scanned using the `oysteR` package against the OSS Index vulnerability database.

```{r vuln-scan, results='asis'}
if (nrow(vuln_report) > 0) {
  vuln_display <- vuln_report
  # Truncate any long description fields for display
  for (col in names(vuln_display)) {
    if (is.character(vuln_display[[col]])) {
      vuln_display[[col]] <- substr(vuln_display[[col]], 1, 60)
    }
  }
  kable(vuln_display, format = "latex", booktabs = TRUE, longtable = TRUE,
        row.names = FALSE, linesep = "")
} else {
  cat("\\textbf{No vulnerabilities detected.} All installed packages passed the OSS Index scan.")
}
```

\newpage

# Reproducibility Manifest

Package sources and exact commit SHAs for audit trail and reproduction purposes.
A `renv.lock` file capturing all installed package versions is provided as a companion artifact.

```{r manifest}
manifest_rows <- lapply(classification$intended_for_use, function(pkg) {
  v <- validation[[pkg]]
  data.frame(
    Package = pkg,
    Source  = trunc_str(pkg_source(pkg, v), 55),
    Git_SHA = if (!is.null(v$git_sha) && !is.na(v$git_sha))
                substr(v$git_sha, 1, 12) else "N/A",
    stringsAsFactors = FALSE
  )
})

if (length(manifest_rows) > 0) {
  manifest_df <- do.call(rbind, manifest_rows)
  kable(manifest_df, format = "latex", booktabs = TRUE,
        row.names = FALSE,
        col.names = c("Package", "Source", "Git SHA (short)"),
        linesep = "") |>
    kableExtra::column_spec(1, width = "4cm") |>
    kableExtra::column_spec(2, width = "7cm") |>
    kableExtra::column_spec(3, width = "3cm")
} else {
  cat("No manifest entries.")
}
```

\newpage

# Appendix A: Session Information {-}

```{r session-info, results='asis'}
session_file <- "session-info.txt"
if (file.exists(session_file)) {
  si_lines <- readLines(session_file)
  cat("\\begin{verbatim}\n")
  cat(paste(si_lines, collapse = "\n"))
  cat("\n\\end{verbatim}\n")
} else {
  si <- sessionInfo()
  cat("\\begin{verbatim}\n")
  cat(paste(capture.output(print(si)), collapse = "\n"))
  cat("\n\\end{verbatim}\n")
}
```
